---
title: "Diamond Signal"
toc: false
---

```{r}
#| echo: false
#| results: asis
source("R/site_helpers.R")
status <- read_site_csv("pipeline_status.csv")
summary <- read_site_csv("model_summary.csv")
board <- read_site_csv("latest_board.csv")
is_live <- nrow(status) && identical(status$line_feed_status[1], "Live")
```

<section class="hero">
<div class="eyebrow">MLB prop intelligence · evidence first</div>
<h1>Find the distance<br>between <span class="accent">price</span> and probability.</h1>
<p class="hero-copy">Diamond Signal turns player form, opponent context, venue conditions, and live sportsbook prices into one readable question: where does the model disagree with the market enough to matter?</p>
<div class="hero-actions">
<a class="button-primary" href="today.html">Open the model board</a>
<a class="button-ghost" href="results.html">See how it works</a>
</div>
</section>

```{r}
#| echo: false
#| results: asis
dot_class <- if (is_live) "status-dot live" else "status-dot"
feed_text <- if (is_live) {
  paste0("Live card captured ", escape_html(status$last_line_refresh[1]))
} else {
  "Model engine ready · live line feed awaiting an Odds API key or available markets"
}
cat(sprintf('<div class="status-strip"><span class="%s"></span><strong>%s</strong></div>', dot_class, feed_text))
```

```{r}
#| echo: false
#| results: asis
if (nrow(summary)) {
  tb <- summary[summary$market == "batter_total_bases", , drop = FALSE]
  ks <- summary[summary$market == "pitcher_strikeouts", , drop = FALSE]
  nrfi <- summary[summary$market == "nrfi", , drop = FALSE]
  cards <- c(
    metric_card("Markets modeled", "3", "Total bases · strikeouts · NRFI", "Each market has its own evaluated model."),
    metric_card("Holdout predictions", format(sum(summary$test_rows), big.mark = ","), "Later games only", "Models are evaluated on games after their training window."),
    metric_card("Best live edge", if (nrow(board)) paste0("+", pct(max(board$probability_difference))) else "Waiting", if (nrow(board)) "Current priced card" else "No line snapshot yet", "Edge is model probability minus no-vig market probability.")
  )
  cat('<div class="metric-grid">', paste(cards, collapse = ""), '</div>')
}
```

## Three markets. One decision language.

<div class="market-grid">
<div class="market-card" style="--accent:#72e0c4"><span class="card-kicker">Player prop</span><h3>Pitcher strikeouts</h3><p>Projects a starter's expected strikeouts from recent form, workload, location, and opponent context, then converts that count into over/under probabilities.</p></div>
<div class="market-card" style="--accent:#71b7ff"><span class="card-kicker">Player prop</span><h3>Batter total bases</h3><p>Estimates total bases using recent production and the environment around the matchup. Context earns a place only when it improves later-game testing.</p></div>
<div class="market-card" style="--accent:#c8f169"><span class="card-kicker">Game market</span><h3>NRFI</h3><p>Estimates the chance of a scoreless first inning from probable starters, recent team NRFI rates, and venue history.</p></div>
</div>

## Built to clarify, not to decorate a guess.

<div class="feature-grid">
<div class="feature-card"><span class="card-kicker">01 · Price aware</span><h3>The line is part of the decision.</h3><p>A projection is not a pick by itself. The board compares modeled probability with the market's no-vig probability at the captured price.</p></div>
<div class="feature-card"><span class="card-kicker">02 · Time ordered</span><h3>Later games test earlier learning.</h3><p>Model evaluation uses chronological holdouts, which is closer to the daily prediction problem than a random split.</p></div>
<div class="feature-card"><span class="card-kicker">03 · Honest states</span><h3>No line, no pick.</h3><p>When data is missing or a model does not clear the edge threshold, the site says so. Empty boards are valid outcomes.</p></div>
</div>
