---
title: "Model Board"
toc: false
---

```{r}
#| echo: false
#| results: asis
source("R/site_helpers.R")
board <- read_site_csv("latest_board.csv")
projections <- read_site_csv("today_projections.csv")
status <- read_site_csv("pipeline_status.csv")
slate_label <- if (nrow(projections) && "slate_label" %in% names(projections)) projections$slate_label[1] else "today"
updated_at <- if (nrow(status) && "site_data_generated_at" %in% names(status)) format_status_time(status$site_data_generated_at[1]) else "Not available"
if (nrow(board)) {
  board_join <- transform(
    board,
    join_market = normalize_market_key(market),
    join_side = ifelse(normalize_market_key(market) == "nrfi", selection, side)
  )
  projections$row_id <- seq_len(nrow(projections))
  projections$join_id <- paste(projections$selection, projections$matchup, projections$market, projections$recommended_side, sep = "||")
  board_join$join_id <- paste(board_join$selection, board_join$matchup, board_join$join_market, board_join$join_side, sep = "||")
  line_lookup <- board_join[!duplicated(board_join$join_id), c("join_id", "american_price")]
  projections <- merge(projections, line_lookup, by = "join_id", all.x = TRUE, sort = FALSE)
  projections <- projections[order(projections$row_id), ]
  projections$join_id <- NULL
  projections$row_id <- NULL
}
```

<section class="page-lead">
<div class="eyebrow">Updated `r updated_at` · `r slate_label` slate</div>
<h1>The model board.</h1>
<p>Use the FanDuel-priced section first when it is available. The model-only board below is the full research slate, with a board score used only to rank ideas against each other, not to claim certainty.</p>
</section>

```{r}
#| echo: false
#| results: asis
captured <- if (nrow(status)) format_status_time(status$last_line_refresh[1]) else "Not available"
qualified <- if (nrow(board)) nrow(board) else 0
cards <- c(
  metric_card("Last site update", updated_at, "Includes board, grading, and site data", "This timestamp should move forward on each refresh."),
  metric_card("Slate date", slate_label, "Games shown on this board", "This date comes from the generated MLB schedule snapshot."),
  metric_card("Priced edges", as.character(qualified), if (qualified) paste("FanDuel snapshot captured", captured) else "No FanDuel edges cleared the current threshold", "This count changes with available lines and the edge threshold.")
)
cat('<div class="metric-grid">', paste(cards, collapse = ""), '</div>')
if (nrow(board)) {
  cat('<h2>FanDuel-priced edges</h2>')
  cat(sprintf('<p class="mini-note">These are the current FanDuel picks that cleared the model edge threshold in this run. If only %s show up, only %s met the filter at the captured prices.</p>', qualified, qualified))
  cat(render_signal_table(board))
}
cat('<h2>Model-only projections</h2>')
cat('<p class="mini-note">This is the broader research board for the slate. FanDuel prices are auto-filled on add when a matching live price exists; otherwise you can still edit the odds manually in My Picks.</p>')
cat(render_projection_board(projections))
```

<div class="callout-panel">
<span class="card-kicker">How to read this board</span>
<h3>Use priced edges as the actionable shortlist.</h3>
<p><strong>FanDuel-priced edges</strong> are the picks where the model currently disagrees most with the captured FanDuel price. <strong>Probability</strong> is the model's estimate for that side. <strong>Score</strong> on the model-only board is just a ranking signal for the slate, not a literal chance of winning.</p>
</div>

## Score guide

<div class="feature-grid">
<div class="feature-card"><span class="pill strong">60+</span><h3>Top of slate</h3><p>Usually the best mix of modeled probability and separation from the reference line. Start here, then confirm price and news.</p></div>
<div class="feature-card"><span class="pill watch">35-60</span><h3>Worth checking</h3><p>Reasonable model support, but more dependent on lineup confirmation, starting pitchers, and what the live FanDuel number is.</p></div>
<div class="feature-card"><span class="pill">Below 35</span><h3>Thin edge</h3><p>Still on the board for transparency, but not something the model is separating strongly from the rest of the slate.</p></div>
</div>
