---
title: "Model Board"
toc: false
---

```{r}
#| echo: false
#| results: asis
source("R/site_helpers.R")
board <- read_site_csv("latest_board.csv")
projections <- read_site_csv("today_projections.csv")
status <- read_site_csv("pipeline_status.csv")
slate_label <- if (nrow(projections) && "slate_label" %in% names(projections)) projections$slate_label[1] else "today"
```

<section class="page-lead">
<div class="eyebrow">4 AM ET model snapshot · `r slate_label`</div>
<h1>The model board.</h1>
<p>Explore today's ranked card. Props are ordered by a uniform confidence score that blends the model's selected-side probability with how far the projection is from the displayed line. NRFI rows use the selected side's probability directly.</p>
</section>

```{r}
#| echo: false
#| results: asis
line_status <- if (nrow(status)) status$line_feed_status[1] else "Unknown"
captured <- if (nrow(status)) status$last_line_refresh[1] else "Not available"
qualified <- if (nrow(board)) sum(board$probability_difference >= .03) else 0
cards <- c(
  metric_card("Line feed", line_status, paste("Last refresh:", captured), "The board cannot price an edge without a sportsbook line."),
  metric_card("Slate date", slate_label, "Games shown on this board", "This date comes from the generated MLB schedule snapshot."),
  metric_card("Best confidence", if (nrow(projections)) sprintf("%.1f", max(projections$confidence)) else "-", "0-100 cross-market score", "A blended score from probability and model-versus-line distance.")
)
cat('<div class="metric-grid">', paste(cards, collapse = ""), '</div>')
if (nrow(board)) {
  cat('<h2>Priced edges</h2>')
  cat(render_signal_table(board))
}
cat('<h2>Model-only projections</h2>')
cat('<p class="mini-note">These recreate a 4 AM ET pregame research board using the MLB schedule, probable starters, forecast weather, and current model inputs. Displayed prop lines are analytical reference lines unless a sportsbook line feed is configured.</p>')
cat(render_projection_board(projections))
```

<div class="callout-panel">
<span class="card-kicker">How to read this board</span>
<h3>Confidence is cross-market, not certainty.</h3>
<p><strong>Projection vs Line</strong> shows the model count against the displayed reference line for props. <strong>Probability</strong> is the chance of the listed side. <strong>Confidence</strong> is a 0-100 score that blends probability with model-versus-line distance so strikeouts, total bases, and NRFI can be compared on one board.</p>
</div>

## Signal guide

<div class="feature-grid">
<div class="feature-card"><span class="pill strong">70+</span><h3>Strong research signal</h3><p>Meaningful probability plus a meaningful model-versus-line gap. Still check lineups, starters, and current prices.</p></div>
<div class="feature-card"><span class="pill watch">45-70</span><h3>Watchlist signal</h3><p>Interesting but more sensitive to model error, late news, or an actual sportsbook line moving away.</p></div>
<div class="feature-card"><span class="pill">Below 45</span><h3>Low separation</h3><p>The model is not far enough from the displayed line or probability midpoint to prioritize it.</p></div>
</div>
