---
title: "01 - MLB Data Acquisition, Context, and Target Construction"
output:
  html_document:
    toc: true
    toc_depth: 3
params:
  start_date: "2025-03-27"
  end_date: !r format(Sys.Date() - 1, "%Y-%m-%d")
  refresh_cache: false
---

## Objective

This notebook acquires completed regular-season MLB games, venue/weather
context, and observed targets for the initial markets:

- Pitcher strikeouts: strikeouts recorded by the starting pitcher.
- Batter total bases: `1B + 2 * 2B + 3 * 3B + 4 * HR`.
- NRFI: whether neither team scored in the first inning.

Official game outcomes, reported weather, and stadium metadata come from
MLB's public Stats API. Raw JSON responses are cached locally so rerunning
downstream analysis does not repeatedly query the service.

```{r setup, message=FALSE, warning=FALSE}
library(dplyr)
library(httr2)
library(jsonlite)
library(knitr)
library(lubridate)
library(purrr)
library(readr)
library(tidyr)

options(dplyr.summarise.inform = FALSE)

project_root <- normalizePath(file.path(getwd(), ".."), mustWork = TRUE)
raw_dir <- file.path(project_root, "data", "raw", "mlb_stats_api")
processed_dir <- file.path(project_root, "data", "processed")
dir.create(raw_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(processed_dir, recursive = TRUE, showWarnings = FALSE)

start_date <- as.Date(params$start_date)
end_date <- as.Date(params$end_date)
stopifnot(start_date <= end_date, end_date < Sys.Date())

`%||%` <- function(x, y) if (is.null(x) || length(x) == 0) y else x
num <- function(x) as.numeric(x %||% NA_real_)
int <- function(x) as.integer(x %||% NA_integer_)
chr <- function(x) as.character(x %||% NA_character_)
```

## Download Helpers

Every response is persisted as JSON in `data/raw/`, which is gitignored. Set
`refresh_cache: true` only when a fresh copy is intentionally required.

```{r download-functions}
read_or_get_json <- function(url, cache_file, refresh = FALSE) {
  if (file.exists(cache_file) && !refresh) {
    return(fromJSON(cache_file, simplifyVector = FALSE))
  }

  response <- request(url) |>
    req_user_agent("Baseball-Props research dashboard (R/httr2)") |>
    req_timeout(seconds = 30) |>
    req_retry(max_tries = 3) |>
    req_perform()

  writeBin(resp_body_raw(response), cache_file)
  fromJSON(cache_file, simplifyVector = FALSE)
}

monthly_windows <- function(first_date, last_date) {
  starts <- seq(floor_date(first_date, "month"), last_date, by = "month")
  tibble(
    start = pmax(starts, first_date),
    end = pmin(ceiling_date(starts, "month") - days(1), last_date)
  )
}

fetch_schedule <- function(first_date, last_date, refresh = FALSE) {
  windows <- monthly_windows(first_date, last_date)

  map2_dfr(windows$start, windows$end, function(window_start, window_end) {
    url <- paste0(
      "https://statsapi.mlb.com/api/v1/schedule?sportId=1&gameTypes=R",
      "&hydrate=venue,weather",
      "&startDate=", window_start, "&endDate=", window_end
    )
    cache_file <- file.path(
      raw_dir,
      paste0("schedule_context_", window_start, "_", window_end, ".json")
    )
    payload <- read_or_get_json(url, cache_file, refresh)

    map_dfr(payload$dates %||% list(), function(day) {
      map_dfr(day$games %||% list(), function(game) {
        tibble(
          game_pk = int(game$gamePk),
          game_date = as.Date(day$date),
          game_datetime = as.POSIXct(game$gameDate, format = "%Y-%m-%dT%H:%M:%OSZ", tz = "UTC"),
          game_state = game$status$abstractGameState %||% NA_character_,
          away_team = game$teams$away$team$name %||% NA_character_,
          home_team = game$teams$home$team$name %||% NA_character_,
          venue_id = int(game$venue$id),
          venue_name = chr(game$venue$name),
          condition = chr(game$weather$condition),
          temperature_f = num(game$weather$temp),
          wind_description = chr(game$weather$wind),
          day_night = chr(game$dayNight)
        )
      })
    })
  }) |>
    distinct(game_pk, .keep_all = TRUE) |>
    filter(game_state == "Final") |>
    arrange(game_date, game_pk)
}
```

## Completed Games

```{r schedule}
games <- fetch_schedule(start_date, end_date, params$refresh_cache)

schedule_summary <- games |>
  summarise(
    start_date = min(game_date),
    end_date = max(game_date),
    completed_games = n()
  )

kable(schedule_summary, caption = "Completed regular-season games acquired")
```

## Stadium and Weather Context

MLB reports the game-time condition, temperature, and wind description on its
schedule feed. Venue hydration provides coordinates, surface, roof type, and
outfield dimensions. These fields are appropriate historical pregame context;
future prediction-time weather forecasts are requested separately in the
daily prediction notebook.

```{r context-functions}
fetch_venue <- function(venue_id, refresh = FALSE) {
  cache_file <- file.path(raw_dir, paste0("venue_", venue_id, ".json"))
  url <- paste0(
    "https://statsapi.mlb.com/api/v1/venues?venueIds=", venue_id,
    "&hydrate=location,fieldInfo"
  )
  venue <- read_or_get_json(url, cache_file, refresh)$venues[[1]]
  location <- venue$location %||% list()
  coordinates <- location$defaultCoordinates %||% list()
  field <- venue$fieldInfo %||% list()

  tibble(
    venue_id = int(venue$id),
    latitude = num(coordinates$latitude),
    longitude = num(coordinates$longitude),
    elevation_ft = num(location$elevation),
    turf_type = chr(field$turfType),
    roof_type = chr(field$roofType),
    left_line_ft = num(field$leftLine),
    left_center_ft = num(field$leftCenter),
    center_ft = num(field$center),
    right_center_ft = num(field$rightCenter),
    right_line_ft = num(field$rightLine)
  )
}

parse_wind_speed <- function(wind_description) {
  as.numeric(sub("^([0-9]+).*", "\\1", wind_description))
}
```

```{r acquire-context}
venues <- map_dfr(
  unique(games$venue_id[!is.na(games$venue_id)]),
  fetch_venue,
  refresh = params$refresh_cache
)

game_context <- games |>
  left_join(venues, by = "venue_id") |>
  mutate(
    wind_mph = parse_wind_speed(wind_description),
    is_night = as.integer(day_night == "night"),
    roof_closed_or_fixed = as.integer(roof_type %in% c("Closed", "Fixed")),
    grass_surface = as.integer(turf_type == "Grass")
  )

write_rds(game_context, file.path(processed_dir, "game_context.rds"))
write_csv(game_context, file.path(processed_dir, "game_context.csv"))
```

## Box Scores

The box-score parser keeps one row per player-game appearance, then produces
separate outcomes for the two markets. Pitcher props are restricted to the
starting pitcher because FanDuel pitcher markets require the listed pitcher to
start for action.

```{r boxscore-functions}
parse_team_players <- function(players, side, game) {
  map_dfr(players %||% list(), function(player) {
    batting <- player$stats$batting %||% list()
    pitching <- player$stats$pitching %||% list()
    team_name <- if (side == "home") game$home_team else game$away_team
    opponent <- if (side == "home") game$away_team else game$home_team

    tibble(
      game_pk = game$game_pk,
      game_date = game$game_date,
      player_id = int(player$person$id),
      player_name = player$person$fullName %||% NA_character_,
      team = team_name,
      opponent = opponent,
      is_home = as.integer(side == "home"),
      plate_appearances = int(batting$plateAppearances),
      at_bats = int(batting$atBats),
      hits = int(batting$hits),
      doubles = int(batting$doubles),
      triples = int(batting$triples),
      home_runs = int(batting$homeRuns),
      innings_pitched = num(pitching$inningsPitched),
      strikeouts = int(pitching$strikeOuts),
      games_started = int(pitching$gamesStarted)
    )
  })
}

fetch_boxscore <- function(game, refresh = FALSE) {
  cache_file <- file.path(raw_dir, paste0("boxscore_", game$game_pk, ".json"))
  url <- paste0("https://statsapi.mlb.com/api/v1/game/", game$game_pk, "/boxscore")
  payload <- read_or_get_json(url, cache_file, refresh)

  bind_rows(
    parse_team_players(payload$teams$away$players, "away", game),
    parse_team_players(payload$teams$home$players, "home", game)
  )
}
```

```{r acquire-boxscores}
player_games <- map_dfr(
  split(games, seq_len(nrow(games))),
  fetch_boxscore,
  refresh = params$refresh_cache
)

batter_games <- player_games |>
  filter(!is.na(plate_appearances), plate_appearances > 0) |>
  mutate(
    across(c(hits, doubles, triples, home_runs), ~ replace_na(.x, 0L)),
    singles = hits - doubles - triples - home_runs,
    total_bases = singles + 2L * doubles + 3L * triples + 4L * home_runs
  ) |>
  select(
    game_pk, game_date, player_id, player_name, team, opponent, is_home,
    plate_appearances, at_bats, hits, doubles, triples, home_runs, total_bases
  )

pitcher_games <- player_games |>
  filter(games_started == 1L, !is.na(strikeouts)) |>
  select(
    game_pk, game_date, player_id, player_name, team, opponent, is_home,
    innings_pitched, strikeouts
  )

write_rds(batter_games, file.path(processed_dir, "batter_game_outcomes.rds"))
write_rds(pitcher_games, file.path(processed_dir, "pitcher_game_outcomes.rds"))
write_csv(batter_games, file.path(processed_dir, "batter_game_outcomes.csv"))
write_csv(pitcher_games, file.path(processed_dir, "pitcher_game_outcomes.csv"))
```

## NRFI Outcome

```{r nrfi-functions}
fetch_first_inning <- function(game, refresh = FALSE) {
  cache_file <- file.path(raw_dir, paste0("linescore_", game$game_pk, ".json"))
  url <- paste0("https://statsapi.mlb.com/api/v1/game/", game$game_pk, "/linescore")
  payload <- read_or_get_json(url, cache_file, refresh)
  innings <- payload$innings %||% list()
  if (length(innings) == 0) {
    return(tibble(
      game_pk = game$game_pk,
      first_inning_away_runs = NA_integer_,
      first_inning_home_runs = NA_integer_
    ))
  }
  first <- innings[[1]]

  tibble(
    game_pk = game$game_pk,
    first_inning_away_runs = int(first$away$runs),
    first_inning_home_runs = int(first$home$runs)
  )
}
```

```{r acquire-nrfi}
nrfi_games <- map_dfr(
  split(games, seq_len(nrow(games))),
  fetch_first_inning,
  refresh = params$refresh_cache
) |>
  left_join(game_context, by = "game_pk") |>
  filter(!is.na(first_inning_away_runs), !is.na(first_inning_home_runs)) |>
  mutate(
    first_inning_runs = first_inning_away_runs + first_inning_home_runs,
    nrfi = as.integer(first_inning_runs == 0)
  )

write_rds(nrfi_games, file.path(processed_dir, "nrfi_game_outcomes.rds"))
write_csv(nrfi_games, file.path(processed_dir, "nrfi_game_outcomes.csv"))
```

## Quality Checks

```{r quality}
quality_summary <- bind_rows(
  batter_games |>
    summarise(
      market = "Batter total bases",
      appearances = n(),
      players = n_distinct(player_id),
      missing_target = sum(is.na(total_bases)),
      mean_target = mean(total_bases)
    ),
  pitcher_games |>
    summarise(
      market = "Pitcher strikeouts",
      appearances = n(),
      players = n_distinct(player_id),
      missing_target = sum(is.na(strikeouts)),
      mean_target = mean(strikeouts)
    ),
  nrfi_games |>
    summarise(
      market = "NRFI",
      appearances = n(),
      players = NA_integer_,
      missing_target = sum(is.na(nrfi)),
      mean_target = mean(nrfi)
    )
)

kable(quality_summary, digits = 2, caption = "Outcome dataset checks")

stopifnot(
  nrow(batter_games) > 0,
  nrow(pitcher_games) > 0,
  nrow(nrfi_games) > 0,
  !anyNA(batter_games$total_bases),
  !anyNA(pitcher_games$strikeouts),
  !anyNA(nrfi_games$nrfi),
  all(batter_games$total_bases >= 0),
  all(pitcher_games$strikeouts >= 0),
  all(nrfi_games$nrfi %in% c(0, 1))
)
```

## Output

The processed files written by this notebook are the input to
`02_feature_engineering.Rmd`. They remain local analysis artifacts until the
pipeline is validated and integrated into the Quarto site.
