---
title: "02 - Pregame Feature Engineering"
output:
  html_document:
    toc: true
    toc_depth: 3
---

## Objective

This notebook converts game outcomes into modeling rows. Every predictor is
calculated using games strictly before the target game; the current game's
result never appears in its own inputs.

```{r setup, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(knitr)
library(purrr)
library(readr)
library(tidyr)

project_root <- normalizePath(file.path(getwd(), ".."), mustWork = TRUE)
processed_dir <- file.path(project_root, "data", "processed")
batter_games <- readRDS(file.path(processed_dir, "batter_game_outcomes.rds"))
pitcher_games <- readRDS(file.path(processed_dir, "pitcher_game_outcomes.rds"))

rolling_prior_mean <- function(x, window) {
  map_dbl(seq_along(x), function(i) {
    prior <- x[seq_len(i - 1)]
    if (length(prior) == 0) NA_real_ else mean(tail(prior, window), na.rm = TRUE)
  })
}

rolling_prior_sd <- function(x, window) {
  map_dbl(seq_along(x), function(i) {
    prior <- tail(x[seq_len(i - 1)], window)
    if (length(prior) < 2) NA_real_ else sd(prior, na.rm = TRUE)
  })
}
```

## Market-Level Modeling Frame

```{r combine-markets}
market_games <- bind_rows(
  batter_games |>
    transmute(
      market = "batter_total_bases",
      game_pk, game_date, player_id, player_name, team, opponent, is_home,
      workload = plate_appearances,
      outcome = total_bases
    ),
  pitcher_games |>
    transmute(
      market = "pitcher_strikeouts",
      game_pk, game_date, player_id, player_name, team, opponent, is_home,
      workload = innings_pitched,
      outcome = strikeouts
    )
) |>
  arrange(market, player_id, game_date, game_pk)
```

## Rolling Player and Opponent Form

For each player, form measures use trailing 5- and 10-game windows plus the
full prior sample. Opponent allowed form is calculated separately and joined
back by opponent and date; it is retained for exploration and future opponent
adjustments, while the first benchmark model uses only features available
without resolving the player's daily opponent from the odds payload.

```{r rolling-features}
player_features <- market_games |>
  group_by(market, player_id) |>
  mutate(
    games_prior = row_number() - 1L,
    player_mean_5 = rolling_prior_mean(outcome, 5),
    player_mean_10 = rolling_prior_mean(outcome, 10),
    player_mean_season = rolling_prior_mean(outcome, Inf),
    player_sd_10 = rolling_prior_sd(outcome, 10),
    workload_mean_10 = rolling_prior_mean(workload, 10)
  ) |>
  ungroup()

opponent_features <- market_games |>
  arrange(market, opponent, game_date, game_pk) |>
  group_by(market, opponent) |>
  mutate(opponent_allowed_mean_10 = rolling_prior_mean(outcome, 10)) |>
  ungroup() |>
  select(market, game_pk, player_id, opponent_allowed_mean_10)

modeling_frame <- player_features |>
  left_join(opponent_features, by = c("market", "game_pk", "player_id")) |>
  filter(games_prior >= 5) |>
  mutate(
    player_sd_10 = replace_na(player_sd_10, 0),
    workload_mean_10 = coalesce(workload_mean_10, workload)
  ) |>
  filter(
    is.finite(player_mean_5),
    is.finite(player_mean_10),
    is.finite(player_mean_season),
    is.finite(workload_mean_10)
  )

write_rds(modeling_frame, file.path(processed_dir, "modeling_frame.rds"))
write_csv(modeling_frame, file.path(processed_dir, "modeling_frame.csv"))
```

## Current Player Form for Prediction

Daily prediction requires a feature row representing information available
after each player's most recent completed game.

```{r prediction-features}
summarise_current_form <- function(data) {
  data |>
    group_by(market, player_id, player_name) |>
    summarise(
      team = last(team),
      last_game_date = max(game_date),
      games_prior = n(),
      player_mean_5 = mean(tail(outcome, 5)),
      player_mean_10 = mean(tail(outcome, 10)),
      player_mean_season = mean(outcome),
      player_sd_10 = if (n() < 2) 0 else sd(tail(outcome, 10)),
      workload_mean_10 = mean(tail(workload, 10)),
      .groups = "drop"
    ) |>
    filter(games_prior >= 5)
}

current_player_form <- market_games |>
  arrange(market, player_id, game_date, game_pk) |>
  summarise_current_form()

write_rds(current_player_form, file.path(processed_dir, "current_player_form.rds"))

feature_summary <- modeling_frame |>
  group_by(market) |>
  summarise(
    modeling_rows = n(),
    unique_players = n_distinct(player_id),
    first_game = min(game_date),
    last_game = max(game_date),
    mean_outcome = mean(outcome)
  )

kable(feature_summary, digits = 2, caption = "Leakage-safe modeling frame")
```

```{r form-plot, fig.height=4.5, fig.width=8}
ggplot(modeling_frame, aes(player_mean_10, outcome)) +
  geom_point(alpha = 0.08) +
  geom_smooth(method = "lm", se = FALSE, color = "#1f5f8b") +
  facet_wrap(~ market, scales = "free") +
  labs(
    title = "Trailing form and next-game outcome",
    x = "Prior 10-game mean",
    y = "Observed next-game target"
  ) +
  theme_minimal()
```
