---
title: "03 - Enhanced Modeling and Chronological Evaluation"
output:
  html_document:
    toc: true
    toc_depth: 3
params:
  training_fraction: 0.8
---

## Objective

This notebook compares baseline and context-enhanced Poisson models for each
player count target, then fits a dedicated NRFI probability model. All
validation is chronological: earlier games train each model and later games
assess performance.

```{r setup, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(knitr)
library(purrr)
library(readr)
library(tidyr)

project_root <- normalizePath(file.path(getwd(), ".."), mustWork = TRUE)
processed_dir <- file.path(project_root, "data", "processed")
frame <- readRDS(file.path(processed_dir, "modeling_frame.rds"))
nrfi_frame <- readRDS(file.path(processed_dir, "nrfi_modeling_frame.rds"))
stopifnot(params$training_fraction > 0.5, params$training_fraction < 1)

baseline_formula <- outcome ~ player_mean_5 + player_mean_10 +
  player_mean_season + player_sd_10 + workload_mean_10 + is_home

enhanced_formula <- outcome ~ player_mean_5 + player_mean_10 +
  player_mean_season + player_sd_10 + workload_mean_10 + is_home +
  opponent_allowed_mean_10 + temperature_f + wind_mph + is_night +
  roof_closed_or_fixed + grass_surface + elevation_ft + center_ft

poisson_deviance <- function(observed, predicted) {
  component <- ifelse(
    observed == 0,
    predicted,
    observed * log(observed / predicted) - (observed - predicted)
  )
  mean(2 * component)
}

fit_market <- function(market_frame, formula, model_name) {
  game_dates <- sort(unique(market_frame$game_date))
  cutoff_index <- floor(length(game_dates) * params$training_fraction)
  cutoff_date <- game_dates[cutoff_index]
  train <- filter(market_frame, game_date <= cutoff_date)
  test <- filter(market_frame, game_date > cutoff_date)

  model <- glm(formula, family = poisson(link = "log"), data = train)
  predictions <- test |>
    mutate(expected_count = pmax(predict(model, newdata = test, type = "response"), 0.001))

  list(
    model = model, train = train, test = predictions,
    cutoff_date = cutoff_date, model_name = model_name
  )
}

market_frames <- split(frame, frame$market)
baseline_models <- map(market_frames, fit_market, formula = baseline_formula, model_name = "Baseline")
enhanced_models <- map(market_frames, fit_market, formula = enhanced_formula, model_name = "Context enhanced")
market_models <- enhanced_models
```

## Player Prop Count Performance

```{r metrics}
summarise_count_metrics <- function(models) {
  imap_dfr(models, function(result, market_name) {
  result$test |>
    summarise(
      market = market_name,
      model = result$model_name,
      training_rows = nrow(result$train),
      test_rows = n(),
      training_through = as.character(result$cutoff_date),
      test_start = as.character(min(game_date)),
      rmse = sqrt(mean((outcome - expected_count)^2)),
      mae = mean(abs(outcome - expected_count)),
      poisson_deviance = poisson_deviance(outcome, expected_count),
      predicted_average = mean(expected_count),
      observed_average = mean(outcome)
    )
  })
}

count_metrics <- bind_rows(
  summarise_count_metrics(baseline_models),
  summarise_count_metrics(enhanced_models)
)

kable(count_metrics, digits = 3, caption = "Chronological holdout count metrics")
```

## Over Probability Evaluation

Prop recommendations need probabilities, not just expected counts. Each
enhanced-model held-out outcome is evaluated against common half-point lines.
This assesses probability calibration without implying historical line
availability.

```{r probability-evaluation}
line_grid <- list(
  batter_total_bases = c(0.5, 1.5, 2.5, 3.5),
  pitcher_strikeouts = c(3.5, 4.5, 5.5, 6.5, 7.5, 8.5)
)

make_line_predictions <- function(results) {
  imap_dfr(results, function(result, market_name) {
    tidyr::crossing(result$test, line = line_grid[[market_name]]) |>
      mutate(
        model = result$model_name,
        over_result = as.integer(outcome > line),
        probability_over = ppois(line, expected_count, lower.tail = FALSE)
      )
  })
}

line_predictions <- bind_rows(
  make_line_predictions(baseline_models),
  make_line_predictions(enhanced_models)
)

probability_metrics <- line_predictions |>
  group_by(market, model) |>
  summarise(
    line_predictions = n(),
    brier_score = mean((over_result - probability_over)^2),
    log_loss = -mean(
      over_result * log(pmax(probability_over, 1e-8)) +
        (1 - over_result) * log(pmax(1 - probability_over, 1e-8))
    )
  )

kable(probability_metrics, digits = 4, caption = "Held-out over probability metrics")

selected_prop_models <- probability_metrics |>
  group_by(market) |>
  slice_min(brier_score, n = 1, with_ties = FALSE) |>
  ungroup() |>
  select(market, model)

calibration <- line_predictions |>
  inner_join(selected_prop_models, by = c("market", "model")) |>
  mutate(probability_bin = cut(
    probability_over,
    breaks = seq(0, 1, by = 0.1),
    include.lowest = TRUE
  )) |>
  group_by(market, model, probability_bin) |>
  summarise(
    predictions = n(),
    predicted = mean(probability_over),
    observed = mean(over_result),
    .groups = "drop"
  )
```

```{r calibration-plot, fig.height=4.5, fig.width=8}
ggplot(calibration, aes(predicted, observed, size = predictions)) +
  geom_abline(linetype = "dashed", color = "gray45") +
  geom_point(color = "#1f5f8b", alpha = 0.85) +
  facet_wrap(~ market) +
  coord_equal(xlim = c(0, 1), ylim = c(0, 1)) +
  labs(
    title = "Chronological holdout calibration",
    x = "Modeled over probability",
    y = "Observed over frequency",
    size = "Rows"
  ) +
  theme_minimal()
```

## NRFI Probability Model

NRFI is a binary outcome, so it uses logistic regression. Its context model
uses both expected starters' prior form, club and venue first-inning history,
and the venue/weather conditions already measured for player props.

```{r nrfi-model}
nrfi_baseline_formula <- nrfi ~ home_player_mean_10 + away_player_mean_10 +
  home_workload_mean_10 + away_workload_mean_10 +
  home_nrfi_mean_10 + away_nrfi_mean_10 + venue_nrfi_mean_20

nrfi_enhanced_formula <- update(
  nrfi_baseline_formula,
  . ~ . + temperature_f + wind_mph + is_night +
    roof_closed_or_fixed + grass_surface + elevation_ft + center_ft
)

fit_nrfi <- function(formula, model_name) {
  dates <- sort(unique(nrfi_frame$game_date))
  cutoff <- dates[floor(length(dates) * params$training_fraction)]
  train <- filter(nrfi_frame, game_date <= cutoff)
  test <- filter(nrfi_frame, game_date > cutoff)
  model <- glm(formula, family = binomial(link = "logit"), data = train)
  test <- test |>
    mutate(probability_nrfi = predict(model, newdata = test, type = "response"))
  list(model = model, train = train, test = test, cutoff_date = cutoff, model_name = model_name)
}

binary_auc <- function(observed, probability) {
  positives <- probability[observed == 1]
  negatives <- probability[observed == 0]
  comparisons <- outer(positives, negatives, "-")
  mean(comparisons > 0) + 0.5 * mean(comparisons == 0)
}

summarise_nrfi <- function(result) {
  result$test |>
    summarise(
      market = "nrfi",
      model = result$model_name,
      training_rows = nrow(result$train),
      test_rows = n(),
      training_through = as.character(result$cutoff_date),
      test_start = as.character(min(game_date)),
      accuracy_at_50 = mean((probability_nrfi >= 0.5) == nrfi),
      brier_score = mean((nrfi - probability_nrfi)^2),
      log_loss = -mean(
        nrfi * log(pmax(probability_nrfi, 1e-8)) +
          (1 - nrfi) * log(pmax(1 - probability_nrfi, 1e-8))
      ),
      auc = binary_auc(nrfi, probability_nrfi),
      predicted_rate = mean(probability_nrfi),
      observed_rate = mean(nrfi)
    )
}

nrfi_baseline <- fit_nrfi(nrfi_baseline_formula, "Baseline")
nrfi_enhanced <- fit_nrfi(nrfi_enhanced_formula, "Context enhanced")
nrfi_constant <- list(
  model_name = "Constant training rate",
  cutoff_date = nrfi_baseline$cutoff_date,
  train = nrfi_baseline$train,
  test = nrfi_baseline$test |>
    mutate(probability_nrfi = mean(nrfi_baseline$train$nrfi))
)
nrfi_metrics <- bind_rows(
  summarise_nrfi(nrfi_constant),
  summarise_nrfi(nrfi_baseline),
  summarise_nrfi(nrfi_enhanced)
)

kable(nrfi_metrics, digits = 4, caption = "Chronological holdout NRFI metrics")
```

## Persist Models and Evaluation

```{r output}
models <- imap(selected_prop_models$market, function(market_name, index) {
  candidate <- selected_prop_models$model[[index]]
  if (candidate == "Context enhanced") {
    enhanced_models[[market_name]]$model
  } else {
    baseline_models[[market_name]]$model
  }
}) |>
  set_names(selected_prop_models$market)
selected_prop_results <- imap(selected_prop_models$market, function(market_name, index) {
  if (selected_prop_models$model[[index]] == "Context enhanced") {
    enhanced_models[[market_name]]
  } else {
    baseline_models[[market_name]]
  }
}) |>
  set_names(selected_prop_models$market)

selected_nrfi_name <- nrfi_metrics |>
  filter(model != "Constant training rate") |>
  slice_min(brier_score, n = 1, with_ties = FALSE) |>
  pull(model)
selected_nrfi_model <- if (selected_nrfi_name == "Context enhanced") {
  nrfi_enhanced$model
} else {
  nrfi_baseline$model
}
selected_nrfi_result <- if (selected_nrfi_name == "Context enhanced") {
  nrfi_enhanced
} else {
  nrfi_baseline
}

holdout_predictions <- map_dfr(selected_prop_results, "test", .id = "market")
selected_registry <- bind_rows(
  selected_prop_models,
  tibble(market = "nrfi", model = selected_nrfi_name)
)

saveRDS(models, file.path(processed_dir, "benchmark_models.rds"))
saveRDS(selected_nrfi_model, file.path(processed_dir, "nrfi_model.rds"))
write_csv(count_metrics, file.path(processed_dir, "count_metrics.csv"))
write_csv(probability_metrics, file.path(processed_dir, "probability_metrics.csv"))
write_csv(nrfi_metrics, file.path(processed_dir, "nrfi_metrics.csv"))
write_csv(holdout_predictions, file.path(processed_dir, "holdout_predictions.csv"))
write_csv(selected_nrfi_result$test, file.path(processed_dir, "nrfi_holdout_predictions.csv"))
write_csv(selected_registry, file.path(processed_dir, "selected_model_registry.csv"))

cat(
  "Candidate models selected by lower holdout Brier score:\n"
)
print(selected_registry)
cat(
  "\nThis report measures predictive accuracy and calibration. Daily captured\n",
  "lines will later support transparent grading of issued recommendations.\n"
)
```
