---
title: "03 - Modeling and Chronological Evaluation"
output:
  html_document:
    toc: true
    toc_depth: 3
params:
  training_fraction: 0.8
---

## Objective

The initial benchmark is a separate Poisson regression for each count target.
It is deliberately interpretable and establishes an honest baseline before
more flexible models are considered. Validation is chronological: earlier
games train the model and later games assess performance.

```{r setup, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(knitr)
library(purrr)
library(readr)
library(tidyr)

project_root <- normalizePath(file.path(getwd(), ".."), mustWork = TRUE)
processed_dir <- file.path(project_root, "data", "processed")
frame <- readRDS(file.path(processed_dir, "modeling_frame.rds"))
stopifnot(params$training_fraction > 0.5, params$training_fraction < 1)

model_formula <- outcome ~ player_mean_5 + player_mean_10 +
  player_mean_season + player_sd_10 + workload_mean_10 + is_home

poisson_deviance <- function(observed, predicted) {
  component <- ifelse(
    observed == 0,
    predicted,
    observed * log(observed / predicted) - (observed - predicted)
  )
  mean(2 * component)
}

fit_market <- function(market_frame) {
  game_dates <- sort(unique(market_frame$game_date))
  cutoff_index <- floor(length(game_dates) * params$training_fraction)
  cutoff_date <- game_dates[cutoff_index]
  train <- filter(market_frame, game_date <= cutoff_date)
  test <- filter(market_frame, game_date > cutoff_date)

  model <- glm(model_formula, family = poisson(link = "log"), data = train)
  predictions <- test |>
    mutate(expected_count = pmax(predict(model, newdata = test, type = "response"), 0.001))

  list(model = model, train = train, test = predictions, cutoff_date = cutoff_date)
}

market_models <- split(frame, frame$market) |>
  map(fit_market)
```

## Count Prediction Performance

```{r metrics}
count_metrics <- imap_dfr(market_models, function(result, market_name) {
  result$test |>
    summarise(
      market = market_name,
      training_rows = nrow(result$train),
      test_rows = n(),
      training_through = as.character(result$cutoff_date),
      test_start = as.character(min(game_date)),
      rmse = sqrt(mean((outcome - expected_count)^2)),
      mae = mean(abs(outcome - expected_count)),
      poisson_deviance = poisson_deviance(outcome, expected_count),
      predicted_average = mean(expected_count),
      observed_average = mean(outcome)
    )
})

kable(count_metrics, digits = 3, caption = "Chronological holdout count metrics")
```

## Over Probability Evaluation

Prop recommendations need probabilities, not just expected counts. For an
initial probability check, each held-out outcome is evaluated against common
half-point lines. This assesses calibration without implying that historical
prices were available or that ROI has been backtested.

```{r probability-evaluation}
line_grid <- list(
  batter_total_bases = c(0.5, 1.5, 2.5, 3.5),
  pitcher_strikeouts = c(3.5, 4.5, 5.5, 6.5, 7.5, 8.5)
)

line_predictions <- imap_dfr(market_models, function(result, market_name) {
  tidyr::crossing(result$test, line = line_grid[[market_name]]) |>
    mutate(
      over_result = as.integer(outcome > line),
      probability_over = ppois(line, expected_count, lower.tail = FALSE)
    )
})

probability_metrics <- line_predictions |>
  group_by(market) |>
  summarise(
    line_predictions = n(),
    brier_score = mean((over_result - probability_over)^2),
    log_loss = -mean(
      over_result * log(pmax(probability_over, 1e-8)) +
        (1 - over_result) * log(pmax(1 - probability_over, 1e-8))
    )
  )

kable(probability_metrics, digits = 4, caption = "Held-out over probability metrics")

calibration <- line_predictions |>
  mutate(probability_bin = cut(
    probability_over,
    breaks = seq(0, 1, by = 0.1),
    include.lowest = TRUE
  )) |>
  group_by(market, probability_bin) |>
  summarise(
    predictions = n(),
    predicted = mean(probability_over),
    observed = mean(over_result),
    .groups = "drop"
  )
```

```{r calibration-plot, fig.height=4.5, fig.width=8}
ggplot(calibration, aes(predicted, observed, size = predictions)) +
  geom_abline(linetype = "dashed", color = "gray45") +
  geom_point(color = "#1f5f8b", alpha = 0.85) +
  facet_wrap(~ market) +
  coord_equal(xlim = c(0, 1), ylim = c(0, 1)) +
  labs(
    title = "Chronological holdout calibration",
    x = "Modeled over probability",
    y = "Observed over frequency",
    size = "Rows"
  ) +
  theme_minimal()
```

## Persist Models and Evaluation

```{r output}
models <- map(market_models, "model")
holdout_predictions <- map_dfr(market_models, "test", .id = "market")

saveRDS(models, file.path(processed_dir, "benchmark_models.rds"))
write_csv(count_metrics, file.path(processed_dir, "count_metrics.csv"))
write_csv(probability_metrics, file.path(processed_dir, "probability_metrics.csv"))
write_csv(holdout_predictions, file.path(processed_dir, "holdout_predictions.csv"))

cat(
  "ROI is intentionally not reported here: valid ROI evaluation requires\n",
  "sportsbook prices captured before each game. That begins in the daily\n",
  "prediction notebook once an odds API credential is configured.\n"
)
```
