---
title: "04 - FanDuel Lines and Daily Predictions"
output:
  html_document:
    toc: true
    toc_depth: 3
params:
  event_date: !r Sys.Date()
  event_timezone: "America/New_York"
  minimum_edge: 0.02
  minimum_ev: 0.03
---

## Objective

This notebook is the bridge from an evaluated model to actionable daily
research. It requests MLB player props from The Odds API, restricts the
sportsbook to FanDuel, attaches model probabilities, and records the exact
price snapshot used to make a recommendation.

It will not publish fabricated picks. If `ODDS_API_KEY` is absent, the
notebook runs through its setup and clearly reports that live scoring is
disabled.

```{r setup, message=FALSE, warning=FALSE}
library(dplyr)
library(httr2)
library(jsonlite)
library(knitr)
library(lubridate)
library(purrr)
library(readr)
library(tidyr)

project_root <- normalizePath(file.path(getwd(), ".."), mustWork = TRUE)
processed_dir <- file.path(project_root, "data", "processed")
prediction_dir <- file.path(project_root, "data", "predictions")
dir.create(prediction_dir, recursive = TRUE, showWarnings = FALSE)

models <- readRDS(file.path(processed_dir, "benchmark_models.rds"))
current_form <- readRDS(file.path(processed_dir, "current_player_form.rds"))
api_key <- Sys.getenv("ODDS_API_KEY")
run_timestamp <- format(Sys.time(), "%Y-%m-%dT%H:%M:%S%z")

if (as.Date(params$event_date) <= max(current_form$last_game_date)) {
  stop("event_date must be later than the outcomes used to build current player form.")
}

clean_name <- function(x) {
  tolower(gsub("[^a-z0-9]", "", iconv(x, to = "ASCII//TRANSLIT")))
}

american_to_decimal <- function(price) {
  ifelse(price > 0, 1 + price / 100, 1 + 100 / abs(price))
}

american_implied_probability <- function(price) {
  ifelse(price > 0, 100 / (price + 100), abs(price) / (abs(price) + 100))
}
```

## Obtain FanDuel Prop Snapshots

The two requested market keys are documented in The Odds API MLB player-prop
catalog: `pitcher_strikeouts` and `batter_total_bases`.

```{r odds-functions}
get_api_json <- function(url, ...) {
  request(url) |>
    req_url_query(apiKey = api_key, ...) |>
    req_user_agent("Baseball-Props research dashboard (R/httr2)") |>
    req_timeout(seconds = 30) |>
    req_retry(max_tries = 3) |>
    req_perform() |>
    resp_body_json(simplifyVector = FALSE)
}

fetch_events <- function() {
  get_api_json("https://api.the-odds-api.com/v4/sports/baseball_mlb/events")
}

fetch_fanduel_props <- function(event) {
  payload <- get_api_json(
    paste0(
      "https://api.the-odds-api.com/v4/sports/baseball_mlb/events/",
      event$id,
      "/odds"
    ),
    regions = "us",
    markets = "pitcher_strikeouts,batter_total_bases",
    bookmakers = "fanduel",
    oddsFormat = "american"
  )

  map_dfr(payload$bookmakers %||% list(), function(book) {
    map_dfr(book$markets %||% list(), function(market) {
      map_dfr(market$outcomes %||% list(), function(outcome) {
        tibble(
          event_id = event$id,
          commence_time = event$commence_time,
          home_team = event$home_team,
          away_team = event$away_team,
          sportsbook = book$key,
          market = market$key,
          player_name = outcome$description,
          side = outcome$name,
          line = as.numeric(outcome$point),
          american_price = as.numeric(outcome$price),
          captured_at = run_timestamp
        )
      })
    })
  })
}
```

```{r live-lines}
`%||%` <- function(x, y) if (is.null(x) || length(x) == 0) y else x

if (!nzchar(api_key)) {
  props <- tibble()
  cat("No ODDS_API_KEY found. Live FanDuel prop retrieval and scoring were skipped.\n")
} else {
  events <- fetch_events()
  selected_events <- keep(events, function(event) {
    local_start <- with_tz(ymd_hms(event$commence_time), params$event_timezone)
    as.Date(local_start) == as.Date(params$event_date)
  })
  props <- map_dfr(selected_events, fetch_fanduel_props)

  if (nrow(props) > 0) {
    write_csv(
      props,
      file.path(
        prediction_dir,
        paste0("fanduel_props_", format(Sys.time(), "%Y%m%d_%H%M%S"), ".csv")
      )
    )
  }
}
```

## Score Available Lines

```{r score-lines}
if (nrow(props) > 0) {
  scoring_data <- props |>
    mutate(player_key = clean_name(player_name)) |>
    left_join(
      current_form |>
        mutate(player_key = clean_name(player_name)) |>
        select(-player_name),
      by = c("market", "player_key")
    ) |>
    filter(!is.na(player_id)) |>
    mutate(is_home = as.integer(team == home_team))

  scored <- scoring_data |>
    group_split(market) |>
    map_dfr(function(market_rows) {
      model <- models[[unique(market_rows$market)]]
      market_rows |>
        mutate(expected_count = predict(model, newdata = market_rows, type = "response"))
    }) |>
    mutate(
      probability_over = ppois(line, expected_count, lower.tail = FALSE),
      model_probability = if_else(side == "Over", probability_over, 1 - probability_over),
      decimal_price = american_to_decimal(american_price),
      implied_probability_raw = american_implied_probability(american_price)
    ) |>
    group_by(event_id, market, player_name, line) |>
    mutate(
      implied_probability = if_else(
        n() == 2L,
        implied_probability_raw / sum(implied_probability_raw),
        implied_probability_raw
      ),
      edge = model_probability - implied_probability,
      expected_value = model_probability * decimal_price - 1,
      qualifies = edge >= params$minimum_edge & expected_value >= params$minimum_ev
    ) |>
    ungroup() |>
    arrange(desc(qualifies), desc(expected_value))

  recommendations <- filter(scored, qualifies)
  write_csv(scored, file.path(prediction_dir, "latest_scored_props.csv"))
  write_csv(recommendations, file.path(prediction_dir, "latest_recommendations.csv"))

  kable(
    scored |>
      select(
        player_name, market, side, line, american_price, expected_count,
        model_probability, implied_probability, edge, expected_value, qualifies
      ),
    digits = 3,
    caption = "Scored FanDuel player props"
  )
} else {
  cat("No prop lines were available to score in this run.\n")
}
```

## Evaluation Discipline

A qualified recommendation becomes eligible for performance reporting only
after this notebook has stored its pregame line and price. Grading future
outcomes against those locked snapshots will support units, ROI, calibration,
and market-by-market performance without look-ahead bias.
