---
title: "04 - FanDuel Lines and Daily Predictions"
output:
  html_document:
    toc: true
    toc_depth: 3
params:
  event_date: !r Sys.Date()
  event_timezone: "America/New_York"
  minimum_probability_difference: 0.03
---

## Objective

This notebook joins evaluated models to daily lines. It requests FanDuel
player props and first-inning totals from The Odds API, retrieves forecast-time
weather for the scheduled MLB venue from Open-Meteo, and stores the exact
snapshot used for each recommendation.

NRFI is represented by the `Under 0.5` outcome in the
`totals_1st_1_innings` market. If `ODDS_API_KEY` is absent, the notebook runs
its setup and reports that live scoring is disabled.

```{r setup, message=FALSE, warning=FALSE}
library(dplyr)
library(httr2)
library(jsonlite)
library(knitr)
library(lubridate)
library(purrr)
library(readr)
library(tidyr)

project_root <- normalizePath(file.path(getwd(), ".."), mustWork = TRUE)
processed_dir <- file.path(project_root, "data", "processed")
prediction_dir <- file.path(project_root, "data", "predictions")
dir.create(prediction_dir, recursive = TRUE, showWarnings = FALSE)

models <- readRDS(file.path(processed_dir, "benchmark_models.rds"))
nrfi_model <- readRDS(file.path(processed_dir, "nrfi_model.rds"))
current_form <- readRDS(file.path(processed_dir, "current_player_form.rds"))
current_opponent_form <- readRDS(file.path(processed_dir, "current_opponent_form.rds"))
current_nrfi_team_form <- readRDS(file.path(processed_dir, "current_nrfi_team_form.rds"))
current_nrfi_venue_form <- readRDS(file.path(processed_dir, "current_nrfi_venue_form.rds"))

api_key <- Sys.getenv("ODDS_API_KEY")
run_timestamp <- format(Sys.time(), "%Y-%m-%dT%H:%M:%S%z")
`%||%` <- function(x, y) if (is.null(x) || length(x) == 0) y else x
num <- function(x) as.numeric(x %||% NA_real_)
int <- function(x) as.integer(x %||% NA_integer_)
chr <- function(x) as.character(x %||% NA_character_)

if (as.Date(params$event_date) <= max(current_form$last_game_date)) {
  stop("event_date must be later than the outcomes used to build current player form.")
}

clean_name <- function(x) {
  tolower(gsub("[^a-z0-9]", "", iconv(x, to = "ASCII//TRANSLIT")))
}

american_implied_probability <- function(price) {
  ifelse(price > 0, 100 / (price + 100), abs(price) / (abs(price) + 100))
}

add_no_vig_probability <- function(data) {
  data |>
    mutate(implied_probability_raw = american_implied_probability(american_price)) |>
    group_by(event_id, market, player_name, line) |>
    mutate(
      implied_probability = if_else(
        n() == 2L,
        implied_probability_raw / sum(implied_probability_raw),
        implied_probability_raw
      )
    ) |>
    ungroup()
}
```

## Data Access Helpers

Historical weather in model training comes from MLB's recorded game weather.
For a future card, Open-Meteo supplies hourly forecast weather at the MLB
venue coordinates available before first pitch.

```{r api-functions}
get_public_json <- function(url, ...) {
  request(url) |>
    req_url_query(...) |>
    req_user_agent("Baseball-Props prediction research dashboard (R/httr2)") |>
    req_timeout(seconds = 30) |>
    req_retry(max_tries = 3) |>
    req_perform() |>
    resp_body_json(simplifyVector = FALSE)
}

get_odds_json <- function(url, ...) {
  get_public_json(url, apiKey = api_key, ...)
}

fetch_events <- function() {
  get_odds_json("https://api.the-odds-api.com/v4/sports/baseball_mlb/events")
}

fetch_fanduel_lines <- function(event) {
  payload <- get_odds_json(
    paste0(
      "https://api.the-odds-api.com/v4/sports/baseball_mlb/events/",
      event$id,
      "/odds"
    ),
    regions = "us",
    markets = "pitcher_strikeouts,batter_total_bases,totals_1st_1_innings",
    bookmakers = "fanduel",
    oddsFormat = "american"
  )

  map_dfr(payload$bookmakers %||% list(), function(book) {
    map_dfr(book$markets %||% list(), function(market) {
      map_dfr(market$outcomes %||% list(), function(outcome) {
        tibble(
          event_id = event$id,
          commence_time = event$commence_time,
          home_team = event$home_team,
          away_team = event$away_team,
          sportsbook = book$key,
          market = market$key,
          player_name = chr(outcome$description),
          side = outcome$name,
          line = as.numeric(outcome$point),
          american_price = as.numeric(outcome$price),
          captured_at = run_timestamp
        )
      })
    })
  })
}

fetch_venue <- function(venue_id) {
  payload <- get_public_json(
    "https://statsapi.mlb.com/api/v1/venues",
    venueIds = venue_id,
    hydrate = "location,fieldInfo"
  )$venues[[1]]
  location <- payload$location %||% list()
  coordinates <- location$defaultCoordinates %||% list()
  field <- payload$fieldInfo %||% list()
  tibble(
    venue_id = int(payload$id),
    latitude = num(coordinates$latitude),
    longitude = num(coordinates$longitude),
    elevation_ft = num(location$elevation),
    roof_closed_or_fixed = as.integer(chr(field$roofType) %in% c("Closed", "Fixed")),
    grass_surface = as.integer(chr(field$turfType) == "Grass"),
    center_ft = num(field$center)
  )
}

fetch_forecast_at_start <- function(latitude, longitude, game_datetime) {
  payload <- get_public_json(
    "https://api.open-meteo.com/v1/forecast",
    latitude = latitude,
    longitude = longitude,
    hourly = "temperature_2m,wind_speed_10m,precipitation_probability",
    temperature_unit = "fahrenheit",
    wind_speed_unit = "mph",
    timezone = params$event_timezone,
    forecast_days = 16
  )
  hourly <- tibble(
    forecast_time = ymd_hm(unlist(payload$hourly$time), tz = params$event_timezone),
    temperature_f = unlist(payload$hourly$temperature_2m),
    wind_mph = unlist(payload$hourly$wind_speed_10m),
    precipitation_probability = unlist(payload$hourly$precipitation_probability)
  )
  start_local <- with_tz(ymd_hms(game_datetime), params$event_timezone)
  hourly[which.min(abs(difftime(hourly$forecast_time, start_local, units = "mins"))), ]
}

fetch_daily_game_context <- function() {
  schedule <- get_public_json(
    "https://statsapi.mlb.com/api/v1/schedule",
    sportId = 1,
    date = as.character(as.Date(params$event_date)),
    hydrate = "venue,probablePitcher"
  )
  games <- map_dfr(schedule$dates %||% list(), function(day) {
    map_dfr(day$games %||% list(), function(game) {
      tibble(
        home_team = chr(game$teams$home$team$name),
        away_team = chr(game$teams$away$team$name),
        home_probable_pitcher = chr(game$teams$home$probablePitcher$fullName),
        away_probable_pitcher = chr(game$teams$away$probablePitcher$fullName),
        commence_time = chr(game$gameDate),
        venue_id = int(game$venue$id),
        is_night = as.integer(chr(game$dayNight) == "night")
      )
    })
  })
  venues <- map_dfr(unique(games$venue_id), fetch_venue)
  games |>
    left_join(venues, by = "venue_id") |>
    mutate(
      weather = pmap(
        list(latitude, longitude, commence_time),
        fetch_forecast_at_start
      )
    ) |>
    unnest(weather)
}
```

## Obtain Daily Line and Weather Snapshots

```{r live-lines}
if (!nzchar(api_key)) {
  lines <- tibble(
    event_id = character(),
    commence_time = character(),
    home_team = character(),
    away_team = character(),
    sportsbook = character(),
    market = character(),
    player_name = character(),
    side = character(),
    line = double(),
    american_price = double(),
    captured_at = character()
  )
  daily_context <- tibble()
  cat("No ODDS_API_KEY found. Live FanDuel line retrieval and scoring were skipped.\n")
} else {
  events <- fetch_events()
  selected_events <- keep(events, function(event) {
    local_start <- with_tz(ymd_hms(event$commence_time), params$event_timezone)
    as.Date(local_start) == as.Date(params$event_date)
  })
  lines <- map_dfr(selected_events, fetch_fanduel_lines)
  daily_context <- fetch_daily_game_context()

  if (nrow(lines) > 0) {
    write_csv(
      lines,
      file.path(
        prediction_dir,
        paste0("fanduel_lines_", format(Sys.time(), "%Y%m%d_%H%M%S"), ".csv")
      )
    )
  }
}
```

## Score Player Props

```{r score-player-props}
player_lines <- lines |>
  filter(market %in% c("pitcher_strikeouts", "batter_total_bases"))

if (nrow(player_lines) > 0) {
  scoring_data <- player_lines |>
    left_join(daily_context, by = c("home_team", "away_team", "commence_time")) |>
    mutate(player_key = clean_name(player_name)) |>
    left_join(
      current_form |>
        mutate(player_key = clean_name(player_name)) |>
        select(-player_name),
      by = c("market", "player_key")
    ) |>
    filter(!is.na(player_id)) |>
    mutate(
      is_home = as.integer(team == home_team),
      opponent = if_else(is_home == 1L, away_team, home_team)
    ) |>
    left_join(current_opponent_form, by = c("market", "opponent"))

  scored_props <- scoring_data |>
    group_split(market) |>
    map_dfr(function(market_rows) {
      model <- models[[unique(market_rows$market)]]
      market_rows |>
        mutate(expected_count = predict(model, newdata = market_rows, type = "response"))
    }) |>
    mutate(
      probability_over = ppois(line, expected_count, lower.tail = FALSE),
      model_probability = if_else(side == "Over", probability_over, 1 - probability_over)
    ) |>
    add_no_vig_probability() |>
    mutate(
      probability_difference = model_probability - implied_probability,
      qualifies = probability_difference >= params$minimum_probability_difference
    ) |>
    arrange(desc(qualifies), desc(probability_difference))

  write_csv(scored_props, file.path(prediction_dir, "latest_scored_player_props.csv"))

  kable(
    scored_props |>
      select(
        player_name, market, side, line, model_probability,
        implied_probability, probability_difference, qualifies
      ),
    digits = 3,
    caption = "Scored FanDuel player props"
  )
} else {
  cat("No player prop lines were available to score in this run.\n")
}
```

## Score NRFI

```{r score-nrfi}
nrfi_lines <- lines |>
  filter(market == "totals_1st_1_innings", line == 0.5)

if (nrow(nrfi_lines) > 0) {
  pitcher_form <- current_form |>
    filter(market == "pitcher_strikeouts") |>
    mutate(pitcher_key = clean_name(player_name))

  nrfi_data <- nrfi_lines |>
    left_join(daily_context, by = c("home_team", "away_team", "commence_time")) |>
    mutate(
      home_pitcher_key = clean_name(home_probable_pitcher),
      away_pitcher_key = clean_name(away_probable_pitcher)
    ) |>
    left_join(
      pitcher_form |>
        transmute(
          home_pitcher_key = pitcher_key,
          home_player_mean_10 = player_mean_10,
          home_workload_mean_10 = workload_mean_10
        ),
      by = "home_pitcher_key"
    ) |>
    left_join(
      pitcher_form |>
        transmute(
          away_pitcher_key = pitcher_key,
          away_player_mean_10 = player_mean_10,
          away_workload_mean_10 = workload_mean_10
        ),
      by = "away_pitcher_key"
    ) |>
    left_join(
      current_nrfi_team_form |>
        rename(home_team = team, home_nrfi_mean_10 = nrfi_mean_10),
      by = "home_team"
    ) |>
    left_join(
      current_nrfi_team_form |>
        rename(away_team = team, away_nrfi_mean_10 = nrfi_mean_10),
      by = "away_team"
    ) |>
    left_join(current_nrfi_venue_form, by = "venue_id") |>
    filter(
      !is.na(home_player_mean_10), !is.na(away_player_mean_10),
      !is.na(venue_nrfi_mean_20)
    )

  scored_nrfi <- nrfi_data |>
    mutate(
      probability_nrfi = predict(nrfi_model, newdata = nrfi_data, type = "response"),
      model_probability = if_else(side == "Under", probability_nrfi, 1 - probability_nrfi)
    ) |>
    add_no_vig_probability() |>
    mutate(
      probability_difference = model_probability - implied_probability,
      qualifies = probability_difference >= params$minimum_probability_difference
    ) |>
    arrange(desc(qualifies), desc(probability_difference))

  write_csv(scored_nrfi, file.path(prediction_dir, "latest_scored_nrfi.csv"))

  kable(
    scored_nrfi |>
      select(
        home_team, away_team, side, line, model_probability,
        implied_probability, probability_difference, qualifies
      ),
    digits = 3,
    caption = "Scored FanDuel first-inning totals (Under 0.5 = NRFI)"
  )
} else {
  cat("No FanDuel first-inning total at 0.5 was available to score in this run.\n")
}
```

## Evaluation Discipline

A recommendation becomes eligible for performance reporting only after this
notebook stores its pregame line, model probability, and market-implied
probability. Future grading reports accuracy and calibration from those locked
snapshots without look-ahead bias.
