---
title: "05 - Model Readiness and Publication Gates"
output:
  html_document:
    toc: true
    toc_depth: 3
---

## Purpose

Prediction models should not begin issuing public recommendations simply
because they render or because one holdout statistic looks promising. This
notebook records the minimum evidence required before model output is promoted
to the Quarto site.

```{r setup, message=FALSE, warning=FALSE}
library(dplyr)
library(knitr)
library(readr)

project_root <- normalizePath(file.path(getwd(), ".."), mustWork = TRUE)
processed_dir <- file.path(project_root, "data", "processed")

count_metrics <- read_csv(file.path(processed_dir, "count_metrics.csv"), show_col_types = FALSE)
probability_metrics <- read_csv(
  file.path(processed_dir, "probability_metrics.csv"),
  show_col_types = FALSE
)
nrfi_metrics <- read_csv(file.path(processed_dir, "nrfi_metrics.csv"), show_col_types = FALSE)
```

## Current Chronological Holdout

```{r current-results}
kable(count_metrics, digits = 3, caption = "Player prop count-model comparison")
kable(probability_metrics, digits = 4, caption = "Player prop probability metrics")
kable(nrfi_metrics, digits = 4, caption = "NRFI probability-model comparison")
```

## Publication Gates

The website can show model-development results immediately. Daily
recommendations should remain in research mode until all applicable gates
pass.

```{r gates}
gates <- tibble::tribble(
  ~gate, ~minimum_standard, ~reason,
  "Historical coverage", "At least one completed season plus the current season", "Covers player changes and venue variation.",
  "Chronological test window", "At least 30 later game dates held out", "Prevents random splitting from overstating accuracy.",
  "Context improvement", "Context model is no worse than baseline on holdout Brier/log loss", "Extra variables must add evidence, not decoration.",
  "Calibration", "Brier score improves on a constant-rate baseline and probability bins are inspected", "Recommendations depend on probabilities being believable.",
  "Live shadow tracking", "At least 100 locked, pregame observations per market before promotion", "Confirms the live ingestion and name/line joins behave correctly.",
  "Data health", "No missing probable starters or weather context in a recommended row", "A high-confidence result with missing inputs is not trustworthy."
)

kable(gates, caption = "Required gates before publishing daily recommendations")
```

## Interpreting Accuracy

For pitcher strikeouts and total bases, there is no single acceptable
"accuracy" number until the scored line is known; a predicted count is judged
by MAE/RMSE and its derived over/under probability by Brier score, log loss,
and calibration. A practical starting standard is a clearly lower holdout
Brier score than a constant-probability baseline, stable calibration, and a
live shadow sample before publishing recommendations.

For NRFI, raw classification accuracy can be misleading because the majority
class may already win frequently. NRFI should not be activated merely for
exceeding 50% accuracy: its context model should improve Brier score and log
loss against the baseline model and against predicting the historical NRFI
rate for every game.
